### 2015-07-10 narrative/deliverables.tex
spelling, etc.
minor changes
more formatting and fixing.

\section{Deliverables}
The following are the per year project deliverables:

\textbf{Year 1}
\begin{itemize}
\item Develop new techniques for data description that allow users to 
describe the data utility based on their expectations. (ORNL)

\item Assemble and evaluate general techniques for refactoring of data
that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers/ORNL)
\item Investigate the trade-off between ``filing'' and ``piling'' data. 
Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. (UCSC/Sandia)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia/UCSC)
\end{itemize}

\textbf{Year 2}
\begin{itemize}
\item Develop a new type of querying system that allows an application to ask the storage system
about completion timing information. (ORNL)
\item Evaluate the effectiveness and overhead to re-organize data and use the multi-layer approach 
to staging. (Brown/ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime. 
Research runtime tracking and estimation. (Rutgers/ORNL)
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions. (Sandia)
(Sandia/UCSC)
\end{itemize}

\textbf{Year 3}
\begin{itemize}
\item Work with DOE applications, particularly XGC1, GTC, and SPECFM3D, to assess 
the new I/O interface including both APIs and hints. (ORNL)
\item Research techniques that can enhance migration without making the eventual 
lookup of data unbounded, and manage each tier in a scalable fashion. (ORNL)
\item Demonstrate incorporating A\&A and show scalability under both application load as
well as storage system pressures showing that we can tolerate both loads and maintain quality of
service. (Sandia)
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}

>>>>

\section{Deliverables}
The following are the per year project deliverables:
\small

\textbf{Year 1}
\begin{tightItemize}
\item Develop  techniques for data description  to 
describe the data utility based on user expectations. (ORNL)

\item Evaluate general  techniques for refactoring of data
that will be embedded in the system . (ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers)
\item Investigate the trade-off between ``filing'' and ``piling'' data. 
Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. (UCSC)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia)
\end{tightItemize}

\textbf{Year 2}
\begin{tightItemize}
\item Develop a querying system to allow apps to inquire
about completion timing information. (ORNL)
\item Evaluate the effectiveness and overhead to re-organize data and use the multi-layer approach 
to staging. (ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost trade-off, and appropriately place/move data objects at runtime. 
Research runtime tracking and estimation. (Rutgers)
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions. (Sandia)
\item Prototype time bounded search approach for finding data within the storage system (UCSC)
\end{tightItemize}
\textbf{Year 3}
\begin{tightItemize}
\item Work with DOE applications to assess 
the new I/O interface including both APIs and hints. (ORNL, Rutgers)
\item Prototype new QoS methods for storing \& retrieving data in given time bounds. (ORNL, UCSC)
\item Demonstrate admission control and show scalability for  application loads and 
 storage system pressures to maintain quality of
service. (Sandia, UCSC)
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{tightItemize}
\normalsize

### 2015-07-10 narrative/enabling_tech.tex
more formatting and fixing.

\section{Enabling Technologies}
\label{sec:enabling-technologies}
Significant existing work will serve as the foundation upon which this project
will be performed.
%1
At the lowest level is the Lightweight File Systems ({\bf LWFS}) project~\cite{oldfield:lwfs},
which provides the simplest core services for a storage system with additional functionality
for using auxiliary services.
%For example,
%by eliminating consistency checks from the core system, applications already
%managing consistency can eliminate the redundant check enhancing performance.
%
%The current LWFS project phase is investigating new ways for managing storage
%devices. This project, 
{\bf Sirocco} came out of the LWFS project and offers a peer-to-peer style data storage device
mesh with data migrating based on use, required resilience, and available
resources. For example, when writing data initially into Sirocco, required
resilience characteristics are provided prompting Sirocco to ensure the data is
stored on appropriate storage device(s) that meet the resilience requirement.
When a particular device is full, Sirocco determines if it should migrate the
data to a new location or if it is a copy beyond the resilience requirement. If
this copy must still be protected, Sirocco will find a place of sufficient 
resilience to move the data to before deleting the local copy. Further, when
data are requested, a client can ask any Sirocco server and Sirocco will find
the data and return it to the client. If desired, the canonical data location
can be moved on client request to offer better performance. Additionally,
a stored object can have multiple data forks that can
store different versions or aspects of the same data.  For this project, we
will build on this Sirocco foundation for data placement, migration, and the
version storage (via the fork mechanism).

@@

We will also take advantage of the many innovations discovered in the Ceph~\cite{weil:osdi06} storage
system project. 
%
%  In the LWFS project, Sandia sought to demonstrate stripping
%down a file system to the bare minimum features required for any storage
%system. Then, additional functionality and semantics, such as POSIX-style
%directories and consistency requirements, were added in using auxiliary
%services separate from the stack or by layering providing a new interface to
%the storage system. The project's first phase demonstrated that using a core
%object storage system with integrated security providing authentication and
%authorization services, customized storage interfaces could be developed with
%exactly the overhead required for different applications. The second phase of
%LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
%devices for efficient and scalable storage performance.  Sirocco has been in
%process for three years and has developed a usable storage infrastructure.
%
The {\bf Ceph} project (ceph.com) started at UC Santa Cruz in 2004 and was initially
funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an
object-based parallel file system that addresses the well-known metadata
service bottleneck of creating and accessing 1000s of files from a typical 
checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
pulled into the Linux mainline in 2010, the focus of Ceph shifted to its
object-based storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in OpenStack
production-level deployments and was recently selected as the new backend for
Flickr and Yahoo! Cloud Services. In this project we will leverage two of the
key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
(CRUSH)~\cite{weil:sc06} will help make resource identification and
predictable performance scalable, and (2) extensible storage object
classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
new APIs, metadata vocabulary, and QoS into the storage stack.

>>>>

\section{Enabling Technologies}
\label{sec:enabling-technologies}
Significant existing work will serve as the foundation upon which this project
will be performed.
%1
At the lowest level is the Lightweight File Systems ({\bf LWFS}) project~\cite{oldfield:lwfs},
which provides the simplest core services for a storage system with additional functionality
for using auxiliary services.
%For example,
%by eliminating consistency checks from the core system, applications already
%managing consistency can eliminate the redundant check enhancing performance.
%
%The current LWFS project phase is investigating new ways for managing storage
%devices. This project, 
In this project we will leverage {\bf Sirocco}~\cite{sirocco}, which  offers a peer-to-peer style data storage device
mesh with data migrating based on use, required resilience, and available
resources. When writing data initially into Sirocco, required
resilience characteristics are provided prompting Sirocco to ensure the data is
stored on appropriate storage device(s) that meet the resilience requirement.
When a particular device is full, Sirocco determines if it should migrate the
data to a new location or if it is a copy beyond the resilience requirement. Sirocco also allows
a stored object to have multiple data forks that can
store different versions or aspects of the same data.  For this project, we
will build on this Sirocco foundation for data placement, migration, and the
version storage (via the fork mechanism).

@@

We will also take advantage of the many innovations discovered in the {\bf Ceph}~\cite{weil:osdi06} storage
system. 
%
%  In the LWFS project, Sandia sought to demonstrate stripping
%down a file system to the bare minimum features required for any storage
%system. Then, additional functionality and semantics, such as POSIX-style
%directories and consistency requirements, were added in using auxiliary
%services separate from the stack or by layering providing a new interface to
%the storage system. The project's first phase demonstrated that using a core
%object storage system with integrated security providing authentication and
%authorization services, customized storage interfaces could be developed with
%exactly the overhead required for different applications. The second phase of
%LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
%devices for efficient and scalable storage performance.  Sirocco has been in
%process for three years and has developed a usable storage infrastructure.
%
%The {\bf Ceph} project
% (ceph.com) started at UC Santa Cruz in 2004 and was initially
%funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the\
 The project goal was to create an
object-based parallel file system that addresses the well-known metadata
service bottleneck of creating and accessing 1000s of files from a typical 
checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
pulled into the Linux mainline in 2010, the focus of Ceph shifted to its
object-based storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in OpenStack
production-level deployments and was recently selected as the new backend for
Flickr and Yahoo! Cloud Services. In this project we will leverage two of the
key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
(CRUSH)~\cite{weil:sc06} will help make resource identification and
predictable performance scalable, and (2) extensible storage object
classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
new APIs, metadata vocabulary, and QoS into the storage stack.
